Extreme heterogeneity of influenza virus infection in single cells

### 2018-01-13 paper/ResponseToReviews.tex

- The "replicate" 8-hr sample isn't really a replicate since a different MOI was used. Also, the purpose of the differences in treatment for this sample compared with the 8-hr sample are not well explained.

{\color{red}
Good point that we'll address already in first comment.}

-Figure 3 is not effective in supporting the observations made in the text (that most cells have little viral RNA, and that there's a wide variability in the amount of viral RNA). Separate distributions for the viral RNA would be more telling. Log scale would be more appropriate for showing counts per cell.

{\color{black} 
We have added Figure 3-Figure supplement 1, which uses a log scale to show the distribution of the amount of viral mRNA per cell.
We still also retain Figure 3 to show the non-transformed data.
Many cells have no viral mRNA, and it is not possible to show these non-infected cells on a logged scale, so the new combination of Figure 3 and Figure 3-supplement 1 provides both linear and log-scale views.
Simple histograms or kernel density plots of the distribution of viral mRNA are not effective until we have removed non-infected cells (as is done in Figure 4E after setting the thresholds for calling infection) since the distribution is so strongly peaked at the low end.
}

### 2018-01-13 paper/ResponseToReviews.tex

- Lines 186-8: This statement doesn't make much sense since you are talking about variation in mRNA proportion. Global effects such as cell size should not affect proportion.

{\color{black}
This is a good point. 
Since the statement in question is tangential to our central conclusions, we have just removed the lines in question.
}

- Lines 193-8: Is this really remarkable? As the authors point out, there is an extensive body of work demonstrating complementation during IAV infection.

{\color{black} 
We have removed the word ``Remarkably'' and just pointed out that in some cells, co-infection is responsible for providing the full gene complement.
This new text more clearly explains that it has long been inferred that co-infection can complement missing segments.
Our contribution is to directly observe single cells where we can identify specific genes that come from different co-infecting viruses, thereby providing an explicit single-cell characterization of this phenomenon. 
}

### 2018-01-13 paper/paper.tex
addressed minor comment about "Remarkably"

\subsection{Co-infection can provide infected cells with the full complement of viral genes.}
Our sequencing enables us to identify the rare cells that were co-infected with both wild-type and synonymously barcoded viral variants.
Overall, we captured 10 such co-infected cells that had $>$5\% of their mRNA derived from virus (Figure~\ref{fig:coexpression}).
Seven of these 10 cells expressed all eight viral genes.
Remarkably, the majority (4 of 7) of these cells would \emph{not} have expressed all the viral genes in the absence of co-infection, since they have at least one gene exclusively derived from each viral variant.
For instance, the cell with 11.2\% of its mRNA from virus in the upper right of Figure~\ref{fig:coexpression} expresses M only from the wildtype viral variant, and NP and HA only from the synonymously barcoded variant.
Our data therefore provide direct single-cell support for the idea that co-infection can rescue missing viral genes~\citep{Brooke:2013kb,brooke2014influenza,Fonville:2015cg,aguilera2017plaques}.

>>>>

\subsection{Co-infection can provide infected cells with the full complement of viral genes.}
Our sequencing enables us to identify the rare cells that were co-infected with both wild-type and synonymously barcoded viral variants.
Overall, we captured 10 such co-infected cells that had $>$5\% of their mRNA derived from virus (Figure~\ref{fig:coexpression}).
Seven of these 10 cells expressed all eight viral genes.
The majority (4 of 7) of these cells would \emph{not} have expressed all the viral genes in the absence of co-infection, since they have at least one gene exclusively derived from each viral variant.
For instance, the cell with 11.2\% of its mRNA from virus in the upper right of Figure~\ref{fig:coexpression} expresses M only from the wildtype viral variant, and NP and HA only from the synonymously barcoded variant.
Our data therefore provide the first direct single-cell observation of the fact that co-infection can rescue missing viral genes~\citep{Brooke:2013kb,brooke2014influenza,Fonville:2015cg,aguilera2017plaques}.
